Can We Crunch the Social Graph on the Cheap?

### 2011-01-25 doc/hotpar2011/softxmt-hotpar2011.tex
modified:   softxmt-hotpar2011.tex

\documentclass{acm_proc_article-sp}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{alltt}
\begin{document}

@@

\todo{Change to discuss API:

@@

>>>>

\documentclass{acm_proc_article-sp}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{alltt}
\usepackage{comment}
\begin{document}

@@

Our programming model partitions the address space into a shared
global shared address space and per node private address spaces.
Locality may be exploited directly by the programmer in the node
private address spaces, just as in a conventional cache-coherent
multiprocessor node.  In contrast, the global space is explicitly
uncached.  There, every processor can access any word of memory
directly via its unique global address.

Whereas in the PGAS model coherency of global copies is implictly
maintained across the system, we forego caching of globally shared
data.  A node can exploit locality within a chunk of data stored in
the global space only by first copying it into the node's local space.
In this way, a node can access data at random across the system while
incurring no hardware coherence overhead: consistency is enforced by
our mechanisms for explicit global synchronization.

The model promises efficiency subject to the presence of sufficient
concurrency.  Exactly how that concurrency is expressed by the
programmer is language dependent.  Our programming model serves to
augment threaded programming languages so that they become latency
tolerant on large address spaces.  Software multithreading is used to
provide latency tolerance: a computation that is about to execute a
long latency operation such as a remote memory reference initiates the
operation and then yields to the scheduler, which quickly resumes
execution of another computation.  Concurrency is also used -- via
aggregation of requests to a common node destination -- to make more
efficient use of network bandwidth and in some cases reduce bandwidth
demands.

Synchronization on locations in the node private address space is the
same as on conventional systems.  Synchronization on global addresses
works differently: we provide atomic operations such as {\tt
  int\_fetch\_add} as well as operations using full-bits, as on the
Cray XMT.  As is true for other long latency operations,
synchronization latency is tolerated by yielding to the scheduler.
Even non-blocking global atomics yield: atomic operations executed as
x86 instructions sabbotage concurrency by forcing all outstanding
memory references to complete before new ones can be issued.  By
yielding, the core can switch to another computation while the
synchronization is accomplished outside of the pipeline.  In addition,
the implementation can exploit aggregation when concurrency is
abundant to increase the throughput of global synchronization.

Thus, the programming model provides the programmer with options: express
locality within the private space, and it will be exploited to reduce
latency and bandwidth; express concurrency, and it will be used toward
tolerating latency and reducing the cost of bandwidth.

\begin {comment}
\todo{Change to discuss API:

@@

\end{comment}
